---
title: "NPA Banking Analysis: Trends, IBC Impact & Recovery (2012–2025)"
author: "Pratham Kanbargi"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 10,
  fig.height = 6
)
```

---

## 1. Project Overview

Non-Performing Assets (NPAs) are loans where borrowers have stopped making
interest or principal repayments for 90 days or more. High NPA levels signal
stress in the banking system, tie up capital, and constrain credit growth.

This project analyses NPA trends across Indian bank groups from 2012 to 2025,
with three core questions:

- How did NPA levels evolve across Public Sector, Private Sector, and Foreign Banks?
- Did the Insolvency and Bankruptcy Code (IBC), enacted in 2016, produce a
  statistically measurable reduction in NPA ratios?
- Which bank group recovered fastest after the NPA peak (2017–2020)?

**Data sources (Reserve Bank of India):**

- *Statement of NPAs of Scheduled Commercial Banks* — gross NPA data by bank
  and bank group, 2012–2025
- *Gross and Net NPA of Scheduled Commercial Banks* — net NPA and advances
  data by bank group, 2011–2024

---

## 2. Setup

```{r libraries}
# Install packages if not already present
# install.packages(c('tidyverse', 'dplyr', 'skimr', 'janitor', 'readxl'))

library(tidyverse)
library(skimr)
library(dplyr)
library(janitor)
library(readr)
library(readxl)
library(ggplot2)
```

---

## 3. Data Loading

```{r load-data}
NPA_13_25 <- read_excel(
  "/kaggle/input/datasets/compfigggfg/npa-from-2013-2025/NPA 13-25.xlsx"
)

Gross_and_net_NPA <- read_excel(
  "/kaggle/input/datasets/compfigggfg/gross-and-net-npa/Gross and net NPA.xlsx"
)
```

---

## 4. Exploratory Data Analysis

### 4.1 NPA_13_25 Overview

```{r eda-npa1}
cat("=== NPA_13_25 ===\n")
dim(NPA_13_25)
head(NPA_13_25, 10)
names(NPA_13_25)

cat("\nUnique Years:\n")
sort(unique(NPA_13_25$Year))

cat("\nUnique Banks:\n")
sort(unique(NPA_13_25$Banks))
```

### 4.2 Gross_and_net_NPA Overview

The second dataset is a multi-block Excel file — each bank group's data is
stacked vertically with repeated header rows. Raw output is shown here for
transparency.

```{r eda-gn, results='hide'}
cat("=== Gross_and_net_NPA ===\n")
dim(Gross_and_net_NPA)
names(Gross_and_net_NPA)
head(Gross_and_net_NPA, 15)
print(Gross_and_net_NPA, n = Inf, width = Inf)
```

### 4.3 Data Quality Checks — NPA_13_25

```{r quality-npa1}
cat("--- Check numeric columns stored as chr ---\n")
sapply(NPA_13_25, class)

cat("\n--- NA counts per column ---\n")
NPA_13_25 |>
  dplyr::summarise(dplyr::across(dplyr::everything(), ~sum(is.na(.))))

cat("\n--- Blank/empty strings in character columns ---\n")
NPA_13_25 |>
  dplyr::summarise(dplyr::across(
    dplyr::where(is.character),
    ~sum(. == "" | . == "N.A." | grepl("^\\s*$", .), na.rm = TRUE)
  ))

cat("\n--- Sample values: Gross NPA ratio column ---\n")
head(NPA_13_25$`Gross NPAs to Gross Advances Ratio (%)`, 20)
```

### 4.4 Group-Level Row Coverage

```{r group-coverage}
group_labels <- c(
  "ALL SCHEDULED COMMERCIAL BANKS", "PUBLIC SECTOR BANKS",
  "NATIONALISED BANKS", "STATE BANK OF INDIA AND ITS ASSOCIATES",
  "PRIVATE SECTOR BANKS", "FOREIGN BANKS", "SMALL FINANCE BANKS"
)

NPA_13_25 |>
  dplyr::filter(Banks %in% group_labels) |>
  dplyr::count(Banks, Year) |>
  tidyr::pivot_wider(names_from = Year, values_from = n) |>
  print(width = Inf)
```

### 4.5 Banks with Incomplete Year Coverage (Mergers / Closures)

```{r bank-coverage}
bank_year_coverage <- NPA_13_25 |>
  filter(!Banks %in% group_labels) |>
  count(Banks) |>
  filter(n < 14) |>
  arrange(n)

cat("Banks with incomplete year coverage (merged/acquired/new):\n")
print(bank_year_coverage, n = Inf)
```

### 4.6 Year Labelling Check

The `Year` column stores the end year of each fiscal period. Year 2012 means
fiscal year 2011–12 (end-March 2012).

```{r year-check}
cat("--- What does year 2012 represent? ---\n")
NPA_13_25 |>
  filter(Year == 2012, Banks == "ALL SCHEDULED COMMERCIAL BANKS")
```

### 4.7 Gross_and_net_NPA Structure

```{r gn-structure}
cat("--- Last 3 rows (footnote) ---\n")
tail(Gross_and_net_NPA, 3)

cat("\n--- Rows 3–17 (first data block) ---\n")
Gross_and_net_NPA[3:17, ]

cat("\nRow 4 full values (2023-24, All SCBs):\n")
Gross_and_net_NPA[4, ] |> t()

cat("\nRow 3 full values (header row):\n")
Gross_and_net_NPA[3, ] |> t()

cat("\nFootnote text:\n")
cat(Gross_and_net_NPA[[113, 1]])

cat("\n\n--- Old Private Sector Banks block (rows 42–54) ---\n")
Gross_and_net_NPA[42:54, ] |> print(width = Inf)
```

### 4.8 Year Alignment Verification

Confirming that `NPA_13_25$Year` maps to the end year of the fiscal period
in `Gross_and_net_NPA`.

```{r year-alignment}
scb_npa1 <- NPA_13_25 |>
  dplyr::filter(Banks == "ALL SCHEDULED COMMERCIAL BANKS") |>
  dplyr::transmute(
    year_npa1    = Year,
    gross_adv_npa1 = round(as.numeric(`Gross Advances`))
  )

scb_gn <- Gross_and_net_NPA |>
  dplyr::filter(grepl("^\\d{4}-\\d{2}$", `Year (end-March)`)) |>
  dplyr::slice(1:14) |>
  dplyr::transmute(
    fiscal_yr     = `Year (end-March)`,
    year_as_endyr = as.integer(substr(`Year (end-March)`, 1, 4)) + 1L,
    gross_adv_gn  = round(as.numeric(`Gross Advances`))
  )

dplyr::left_join(scb_npa1, scb_gn, by = c("year_npa1" = "year_as_endyr")) |>
  dplyr::select(year_npa1, fiscal_yr, gross_adv_npa1, gross_adv_gn) |>
  print(n = Inf)
```

---

## 5. Data Cleaning & Integration

### 5.1 Step 1 — Extract Group Rows from NPA_13_25

```{r step1}
npa1_groups <- NPA_13_25 |>
  filter(Banks %in% c(
    "ALL SCHEDULED COMMERCIAL BANKS", "NATIONALISED BANKS",
    "PRIVATE SECTOR BANKS", "PUBLIC SECTOR BANKS", "FOREIGN BANKS"
  )) |>
  transmute(
    year  = as.integer(Year),
    bank  = case_when(
      Banks == "ALL SCHEDULED COMMERCIAL BANKS" ~ "All Scheduled Commercial Banks",
      Banks == "NATIONALISED BANKS"             ~ "Nationalised Banks",
      Banks == "PRIVATE SECTOR BANKS"           ~ "Private Sector Banks",
      Banks == "PUBLIC SECTOR BANKS"            ~ "Public Sector Banks",
      Banks == "FOREIGN BANKS"                  ~ "Foreign Banks"
    ),
    gross_advances = as.numeric(`Gross Advances`),
    gross_npa      = as.numeric(`Gross NPAs`),
    gross_npa_pct  = round(as.numeric(`Gross NPAs to Gross Advances Ratio (%)`), 2)
  )

cat("Groups and year coverage:\n")
npa1_groups |>
  count(bank, year) |>
  tidyr::pivot_wider(names_from = year, values_from = n)
```

### 5.2 Step 2 — Parse Gross_and_net_NPA into Labelled Blocks

The raw file is a vertically-stacked multi-block Excel export. Each bank
group's data block is preceded by a section header row. We identify the row
number of each header to assign block labels.

```{r step2}
gn_raw <- Gross_and_net_NPA |>
  rename(
    year_str             = `Year (end-March)`,
    gross_advances2      = `Gross Advances`,
    net_advances         = `Net Advances`,
    gross_npa2           = `Non-Performing Assets (NPAs)`,
    gross_npa_pct_adv    = ...5,
    gross_npa_pct_assets = ...6,
    net_npa              = ...7,
    net_npa_pct_adv      = ...8,
    net_npa_pct_assets   = ...9
  ) |>
  mutate(row_n = row_number())

# Find section boundary row numbers
label_rows <- gn_raw |>
  filter(year_str %in% c(
    "Public Sector Banks", "Old Private Sector Banks",
    "Private Sector Banks *", "Foreign Banks In India",
    "Small Finance Banks"
  )) |>
  select(row_n, year_str)

r_psb  <- label_rows$row_n[label_rows$year_str == "Public Sector Banks"]
r_opvt <- label_rows$row_n[label_rows$year_str == "Old Private Sector Banks"]
r_pvt  <- label_rows$row_n[label_rows$year_str == "Private Sector Banks *"]
r_forb <- label_rows$row_n[label_rows$year_str == "Foreign Banks In India"]
r_sfb  <- label_rows$row_n[label_rows$year_str == "Small Finance Banks"]

gn_net <- gn_raw |>
  mutate(bank = case_when(
    row_n < r_psb                     ~ "All Scheduled Commercial Banks",
    row_n >= r_psb  & row_n < r_opvt ~ "Public Sector Banks",
    row_n >= r_opvt & row_n < r_pvt  ~ "Old Private Sector Banks",
    row_n >= r_pvt  & row_n < r_forb ~ "Private Sector Banks",
    row_n >= r_forb & row_n < r_sfb  ~ "Foreign Banks",
    row_n >= r_sfb                    ~ "Small Finance Banks"
  )) |>
  filter(grepl("^\\d{4}-\\d{2}$", year_str)) |>
  filter(bank %in% c(
    "All Scheduled Commercial Banks", "Public Sector Banks",
    "Private Sector Banks", "Foreign Banks"
  )) |>
  transmute(
    year        = as.integer(substr(year_str, 1, 4)) + 1L,
    bank,
    net_advances = round(as.numeric(net_advances), 2),
    net_npa      = round(as.numeric(net_npa), 2),
    net_npa_pct  = round(as.numeric(net_npa_pct_adv), 2)
  )

cat("Net NPA data coverage:\n")
gn_net |>
  count(bank, year) |>
  tidyr::pivot_wider(names_from = year, values_from = n)
```

### 5.3 Step 3 — Join and Build Final Table

```{r step3}
npa_clean <- npa1_groups |>
  left_join(gn_net, by = c("year", "bank")) |>
  select(year, bank, gross_advances, gross_npa,
         net_advances, net_npa, gross_npa_pct, net_npa_pct) |>
  arrange(desc(year), bank)

cat("Final table dimensions:", dim(npa_clean), "\n\n")
cat("Rows per bank:\n")
npa_clean |> count(bank)

cat("\nNA summary:\n")
npa_clean |> summarise(across(everything(), ~sum(is.na(.))))

cat("\nRows with missing net data:\n")
npa_clean |>
  filter(is.na(net_advances)) |>
  select(year, bank, gross_advances, net_advances)

npa_final <- npa_clean
print(npa_final, n = 56, width = Inf)

write.csv(npa_final, "/kaggle/working/npa_final.csv", row.names = FALSE)
cat("Exported successfully.\n")
```

---

## 6. Preliminary Visualisations

```{r plot-gross-trend}
ggplot(npa_final, aes(x = year, y = gross_npa_pct, color = bank)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Gross NPA % of Gross Advances by Bank Group",
    x = "Year", y = "Gross NPA (%)", color = NULL
  )

ggsave("/kaggle/working/gross_npa_trend.png", width = 10, height = 6, dpi = 150)
```

```{r plot-net-trend}
ggplot(
  npa_final |> dplyr::filter(!is.na(net_npa_pct)),
  aes(x = year, y = net_npa_pct, color = bank)
) +
  geom_line() +
  geom_point() +
  labs(
    title = "Net NPA % of Net Advances by Bank Group",
    x = "Year", y = "Net NPA (%)", color = NULL
  )

ggsave("/kaggle/working/net_npa_trend.png", width = 10, height = 6, dpi = 150)
```

---

## 7. Pre vs Post IBC Analysis

The Insolvency and Bankruptcy Code (IBC) was enacted in May 2016 to provide a
time-bound resolution process for stressed assets. We compare the mean Gross
NPA % in the pre-IBC period (2012–2016) against the post-IBC period (2017–2023)
for each bank group.

```{r ibc-periods}
npa_periods <- npa_final |>
  filter(year >= 2012, year <= 2023) |>
  mutate(period = if_else(
    year <= 2016, "Pre-IBC (2012-16)", "Post-IBC (2017-23)"
  ))

mean_by_period <- npa_periods |>
  group_by(bank, period) |>
  summarise(
    n          = n(),
    mean_gross = round(mean(gross_npa_pct, na.rm = TRUE), 2),
    .groups    = "drop"
  ) |>
  tidyr::pivot_wider(names_from = period, values_from = c(mean_gross, n))

cat("Mean Gross NPA % by Bank Group and Period:\n")
mean_by_period
```

```{r plot-ibc}
ggplot(
  npa_periods |> filter(!is.na(gross_npa_pct)),
  aes(x = year, y = gross_npa_pct, color = bank)
) +
  annotate("rect", xmin = 2012, xmax = 2016.5,
           ymin = -Inf, ymax = Inf, fill = "steelblue", alpha = 0.08) +
  annotate("rect", xmin = 2016.5, xmax = 2023,
           ymin = -Inf, ymax = Inf, fill = "tomato", alpha = 0.08) +
  annotate("text", x = 2014, y = 14.8,
           label = "Pre-IBC", size = 3.5, color = "steelblue4") +
  annotate("text", x = 2020, y = 14.8,
           label = "Post-IBC", size = 3.5, color = "tomato4") +
  geom_vline(xintercept = 2016.5, linetype = "dashed", color = "grey40") +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = 2012:2023) +
  labs(
    title = "Gross NPA % by Bank Group: Pre vs Post IBC (2016)",
    x = "Year", y = "Gross NPA (% of Gross Advances)", color = NULL
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave("/kaggle/working/pre_post_ibc.png", width = 10, height = 6, dpi = 150)
```

### 7.1 Statistical Test — Wilcoxon Rank-Sum

With only 5 pre-IBC and 7 post-IBC observations per group, parametric tests
are inappropriate. The Wilcoxon rank-sum (Mann-Whitney U) test is used instead.
Groups tested: All SCBs, Foreign Banks, Private Sector Banks — the three groups
with complete data in both periods.

```{r wilcoxon}
testable_groups <- c(
  "All Scheduled Commercial Banks", "Foreign Banks", "Private Sector Banks"
)

test_results <- map_dfr(testable_groups, function(grp) {
  pre  <- npa_periods |>
    filter(bank == grp, period == "Pre-IBC (2012-16)") |>
    pull(gross_npa_pct)
  post <- npa_periods |>
    filter(bank == grp, period == "Post-IBC (2017-23)") |>
    pull(gross_npa_pct)

  wt <- wilcox.test(post, pre, alternative = "two.sided", exact = FALSE)

  tibble(
    bank        = grp,
    mean_pre    = round(mean(pre), 2),
    mean_post   = round(mean(post), 2),
    change      = round(mean(post) - mean(pre), 2),
    W_statistic = wt$statistic,
    p_value     = round(wt$p.value, 4)
  )
})

test_results

write.csv(test_results, "/kaggle/working/wilcoxon_test_results.csv",
          row.names = FALSE)
```

> **Interpretation:** A p-value below 0.05 indicates the difference in NPA
> ratios between pre- and post-IBC periods is statistically significant.
> The `change` column shows the direction: negative = improvement post-IBC.

---

## 8. NPA Peak & Recovery Analysis

### 8.1 YoY Change and Peak Identification

```{r peak}
npa_yoy <- npa_final |>
  filter(!is.na(gross_npa_pct)) |>
  arrange(bank, year) |>
  group_by(bank) |>
  mutate(yoy_change = gross_npa_pct - lag(gross_npa_pct)) |>
  ungroup()

peaks <- npa_yoy |>
  filter(year >= 2017, year <= 2021) |>
  group_by(bank) |>
  slice_max(gross_npa_pct, n = 1) |>
  ungroup() |>
  select(bank, peak_year = year, peak_pct = gross_npa_pct)

cat("Peak year and gross NPA % per group:\n")
peaks
```

### 8.2 Recovery Statistics

```{r recovery}
recovery <- npa_yoy |>
  left_join(peaks, by = "bank") |>
  filter(year > peak_year, year <= 2023) |>
  group_by(bank) |>
  summarise(
    peak_year          = first(peak_year),
    peak_pct           = first(peak_pct),
    recovery_end_pct   = last(gross_npa_pct),
    total_decline      = round(first(peak_pct) - last(gross_npa_pct), 2),
    years_of_recovery  = n(),
    avg_annual_decline = round(
      (first(peak_pct) - last(gross_npa_pct)) / n(), 2
    ),
    .groups = "drop"
  ) |>
  arrange(desc(avg_annual_decline))

cat("Recovery stats (post-peak through 2023):\n")
recovery

write.csv(recovery, "/kaggle/working/recovery_stats.csv", row.names = FALSE)
```

### 8.3 YoY Change Plot (Post-Peak)

```{r plot-yoy}
yoy_postpeak <- npa_yoy |>
  left_join(peaks, by = "bank") |>
  filter(year > peak_year, year <= 2023, !is.na(yoy_change)) |>
  mutate(direction = if_else(yoy_change < 0, "Decline", "Rise"))

ggplot(yoy_postpeak, aes(x = year, y = yoy_change, fill = direction)) +
  geom_col() +
  geom_hline(yintercept = 0, linewidth = 0.4) +
  facet_wrap(~bank, scales = "free_x") +
  scale_fill_manual(values = c("Decline" = "steelblue", "Rise" = "tomato")) +
  labs(
    title = "Year-on-Year Change in Gross NPA % (Post-Peak Recovery)",
    x = "Year", y = "YoY Change (percentage points)", fill = NULL
  ) +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

ggsave("/kaggle/working/yoy_recovery.png", width = 12, height = 6, dpi = 150)
```

### 8.4 Indexed Recovery Plot

NPA % rebased to 100 at each group's peak, making the speed of decline
directly comparable across groups.

```{r plot-indexed}
recovery_indexed <- npa_yoy |>
  left_join(peaks, by = "bank") |>
  filter(year >= peak_year, year <= 2023, !is.na(gross_npa_pct)) |>
  group_by(bank) |>
  mutate(
    years_since_peak = year - peak_year,
    indexed_npa      = round((gross_npa_pct / first(gross_npa_pct)) * 100, 1)
  ) |>
  ungroup() |>
  filter(bank != "Nationalised Banks")  # only 1 post-peak data point

ggplot(recovery_indexed,
       aes(x = years_since_peak, y = indexed_npa, color = bank)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2.5) +
  geom_hline(yintercept = 100, linetype = "dashed", color = "grey50") +
  scale_x_continuous(breaks = 0:6) +
  scale_y_continuous(limits = c(0, 105)) +
  labs(
    title    = "Speed of NPA Recovery: Indexed to Peak = 100",
    subtitle = "How much of peak NPA % remains at each year after peak",
    x = "Years Since Peak", y = "NPA % (peak = 100)", color = NULL
  )

ggsave("/kaggle/working/recovery_indexed.png", width = 10, height = 6,
       dpi = 150)
```

---

## 9. Data Quality Audit

Full validation of `npa_final` before Tableau export.

```{r audit}
cat("=== 1. DIMENSIONS & STRUCTURE ===\n")
cat("Rows:", nrow(npa_final), "| Cols:", ncol(npa_final), "\n")
cat("Year range:", min(npa_final$year), "to", max(npa_final$year), "\n\n")

cat("=== 2. MISSING YEAR × BANK COMBINATIONS ===\n")
expand_grid(year = 2012:2025, bank = unique(npa_final$bank)) |>
  left_join(npa_final |> mutate(exists = TRUE), by = c("year", "bank")) |>
  filter(is.na(exists)) |>
  print(n = Inf)

cat("=== 3. net_advances > gross_advances ===\n")
npa_final |>
  filter(!is.na(net_advances), net_advances > gross_advances) |>
  select(year, bank, gross_advances, net_advances, gross_npa, net_npa)

cat("\n=== 4. net_npa > net_advances ===\n")
npa_final |>
  filter(!is.na(net_npa), !is.na(net_advances), net_npa > net_advances) |>
  select(year, bank, net_advances, net_npa)

cat("\n=== 5. gross_npa > gross_advances ===\n")
npa_final |>
  filter(!is.na(gross_npa), gross_npa > gross_advances) |>
  select(year, bank, gross_advances, gross_npa)

cat("\n=== 6. RATIO CONSISTENCY (recomputed vs stored) ===\n")
npa_final |>
  filter(!is.na(gross_npa_pct)) |>
  mutate(
    recomputed_pct = round((gross_npa / gross_advances) * 100, 2),
    diff           = abs(gross_npa_pct - recomputed_pct)
  ) |>
  filter(diff > 0.05) |>
  select(year, bank, gross_npa_pct, recomputed_pct, diff)
```

### 9.1 Foreign Banks Anomaly Investigation

```{r foreign-banks-audit}
cat("=== FOREIGN BANKS: source-level cross-check ===\n")
cat("--- From NPA_13_25 ---\n")
NPA_13_25 |>
  filter(Banks == "FOREIGN BANKS", Year %in% c(2012, 2013, 2021, 2024)) |>
  mutate(across(c(`Gross Advances`, `Gross NPAs`), as.numeric)) |>
  select(Year, `Gross Advances`, `Gross NPAs`,
         `Gross NPAs to Gross Advances Ratio (%)`)

cat("\n--- From Gross_and_net_NPA (Foreign Banks block) ---\n")
Gross_and_net_NPA[83:96, ] |>
  filter(grepl("^\\d{4}-\\d{2}$", `Year (end-March)`)) |>
  filter(`Year (end-March)` %in% c("2011-12", "2012-13", "2020-21", "2023-24")) |>
  select(`Year (end-March)`, `Gross Advances`, `Net Advances`, `...7`)
```

> **Note:** For Foreign Banks in 2012 and 2013, the gross advances figures
> differ between the two RBI source tables, indicating different reporting
> scopes (on-balance-sheet vs total). These rows' net NPA columns are set to
> NA in the final dataset. This is a source-level limitation, not a
> cleaning error.

### 9.2 Year-Join Mismatch Fix

```{r fix-mismatches}
mismatch_keys <- tribble(
  ~year, ~bank,
  2012L, "All Scheduled Commercial Banks",
  2012L, "Foreign Banks",
  2013L, "Foreign Banks"
)

npa_fixed <- npa_final |>
  left_join(mismatch_keys |> mutate(mismatch = TRUE),
            by = c("year", "bank")) |>
  mutate(
    net_advances = if_else(!is.na(mismatch), NA_real_, net_advances),
    net_npa      = if_else(!is.na(mismatch), NA_real_, net_npa),
    net_npa_pct  = if_else(!is.na(mismatch), NA_real_, net_npa_pct)
  ) |>
  select(-mismatch)

cat("Rows affected (should be 3):\n")
npa_fixed |>
  filter(year %in% c(2012, 2013),
         bank %in% c("All Scheduled Commercial Banks", "Foreign Banks")) |>
  select(year, bank, gross_advances, net_advances, net_npa)
```

### 9.3 Add Skeleton Rows for Structural Completeness

Public Sector Banks data starts from 2018 in source 2 (no earlier net NPA
data). Nationalised Banks was discontinued after 2017 following SBI merger.
Skeleton rows with NA values are added to maintain a complete year × bank
grid for Tableau.

```{r skeleton}
skeleton <- bind_rows(
  tibble(year = 2012:2017, bank = "Public Sector Banks"),
  tibble(year = 2018:2025, bank = "Nationalised Banks")
) |>
  mutate(
    gross_advances = NA_real_, gross_npa    = NA_real_,
    net_advances   = NA_real_, net_npa      = NA_real_,
    gross_npa_pct  = NA_real_, net_npa_pct  = NA_real_
  )

npa_fixed <- bind_rows(npa_fixed, skeleton)

cat("New row count (expect 70):", nrow(npa_fixed), "\n")
npa_fixed |> count(bank)
```

### 9.4 Final Checks

```{r final-checks}
npa_final <- npa_fixed |>
  mutate(across(where(is.double), ~round(., 2))) |>
  arrange(desc(year), bank)

cat("=== FINAL CHECKS ===\n\n")
cat("1. Dimensions (expect 70 x 8):", dim(npa_final), "\n\n")

cat("2. net_advances > gross_advances:\n")
npa_final |>
  filter(!is.na(net_advances), net_advances > gross_advances) |>
  select(year, bank, gross_advances, net_advances) |>
  print()

cat("\n3. NA summary per column:\n")
npa_final |>
  summarise(across(everything(), ~sum(is.na(.)))) |>
  print(width = Inf)

cat("\n4. Missing year × bank combinations:\n")
expand_grid(year = 2012:2025, bank = unique(npa_final$bank)) |>
  anti_join(npa_final, by = c("year", "bank")) |>
  nrow() |>
  cat("Missing combinations:", ... = _, "\n")

write.csv(npa_final, "/kaggle/working/npa_final.csv", row.names = FALSE)
cat("CSV re-exported.\n\n")
print(npa_final, n = 70, width = Inf)
```

---

## 10. Tableau Export

Four CSV files are exported for Tableau dashboards.

```{r tableau-export}
npa_final <- read.csv("/kaggle/working/npa_final.csv")

# Export 1: Trend lines
trend_data <- npa_final |>
  filter(!is.na(gross_npa_pct), year <= 2024) |>
  filter(!bank %in% c("Nationalised Banks", "All Scheduled Commercial Banks")) |>
  select(year, bank, gross_npa_pct, net_npa_pct)

write.csv(trend_data, "/kaggle/working/tableau_trend.csv", row.names = FALSE)

# Export 2: Pre/Post IBC comparison
prepost_data <- npa_final |>
  filter(!is.na(gross_npa_pct), year >= 2012, year <= 2023) |>
  filter(bank %in% c("Public Sector Banks", "Private Sector Banks",
                     "Foreign Banks")) |>
  mutate(period = case_when(
    year <= 2016 ~ "Pre-IBC (2012-16)",
    year >= 2020 ~ "Post-IBC Recovery (2020-23)",
    TRUE         ~ "Stress Peak (2017-19)"
  )) |>
  group_by(bank, period) |>
  summarise(mean_gross_npa_pct = round(mean(gross_npa_pct), 2),
            .groups = "drop")

write.csv(prepost_data, "/kaggle/working/tableau_prepost.csv",
          row.names = FALSE)

# Export 3: Indexed recovery
peaks <- tribble(
  ~bank,                  ~peak_year,
  "Public Sector Banks",  2018,
  "Private Sector Banks", 2020,
  "Foreign Banks",        2017
)

recovery_indexed <- npa_final |>
  filter(!is.na(gross_npa_pct)) |>
  inner_join(peaks, by = "bank") |>
  filter(year >= peak_year, year <= 2023) |>
  group_by(bank) |>
  mutate(
    years_since_peak = year - peak_year,
    indexed_npa      = round((gross_npa_pct / first(gross_npa_pct)) * 100, 1)
  ) |>
  ungroup() |>
  select(bank, year, years_since_peak, gross_npa_pct, indexed_npa)

write.csv(recovery_indexed, "/kaggle/working/tableau_recovery.csv",
          row.names = FALSE)

# Export 4: Full clean table
write.csv(
  npa_final |> filter(!is.na(gross_npa_pct)),
  "/kaggle/working/tableau_full.csv",
  row.names = FALSE
)

cat("All 4 Tableau CSVs exported.\n")
```

---

## 11. Key Findings

```{r findings, echo=FALSE}
cat("Summary of key findings from this analysis:\n\n")
cat("1. PEAK NPA LEVELS: Public Sector Banks peaked at ~14.6% in 2018,\n")
cat("   significantly higher than Private Sector Banks (~4.7% in 2020)\n")
cat("   and Foreign Banks (~3.8% in 2017).\n\n")
cat("2. IBC IMPACT: Post-IBC mean NPA ratios declined across all tested\n")
cat("   groups. See test_results for statistical significance.\n\n")
cat("3. RECOVERY SPEED: Foreign Banks recovered fastest in indexed terms,\n")
cat("   followed by Public Sector Banks. See recovery_stats.csv.\n\n")
cat("4. DATA NOTES: 3 rows (Foreign Banks 2012–2013, All SCBs 2012) have\n")
cat("   net NPA columns set to NA due to source-level year mismatches\n")
cat("   between the two RBI tables. Gross NPA data is unaffected.\n")
```

---

*Data sources: Reserve Bank of India. Analysis by Pratham Kanbargi.*
